\Section{Logging in}
The \TIMO\ protocol has some main features that set it apart from other authentication protocols.
We start here with the seeding ideas and deal with improving it later.
%%%
\Subsection{Basic ideas}
A compromised account can be devastating.
It is worth a considerable amount of money when you have a database full of usernames and their passwords, so hackers are always looking for ways to get them.
What if there were no passwords (or equivalents) to be found?
\par
Another thing bothering me was the use of a string of very few characters to identify a single person out of millions.
A traditional `userid' is typically either a short string of lowercase letters and numbers
(e.g. jessica77)
or an email address
(jessica.chastain@hotmail.com).
But this does not uniquely identify anyone, and anyone can misuse this for illicit purposes.
%%%
\Subsubsection{Eliminating passwords}
You still need to be able to authenticate yourself, otherwise you cannot be granted access to your personal account.
That implies that the website needs to store something to facilitate this basic process.
However, it must not be something useable in and of itself; it has to have a counterpart that is in the posession of the individual user.
This counterpart must be different but related.\footnote{Like the locket with Annie's supposed parents, in the film that bears her name.}
\par
The notion of having different keys for user and server is of paramount importance.
It changes the server from a treasure trove (millions of passwords) to a dry desert (millions of useless keys).
\par
The website has a key for each user, which do not change in principle.
The user has a key that is derived from the key the webserver has for that user.
The user keys are derived from the site keys by encrypting it with another key.
\par
If we imagine a random 128-bit value and its \AES\ encryption, we could store that random value as the key for a user in the website's database and give the user its encrypted counterpart.
With these keys we can design a protocol with which the user can prove it has the right key.
This way, what is stored on the website's disks is no longer usable for logging in, as the key to encrypt them is nowhere to be found.
As it turns out, the user can also verify the validity of the website (separate from any PKI scheme) with the same protocol.
%%%
\Subsubsection{The key ring}
Passwords you can remember; keys you can't\footnote{Well, that is true for most people. Some are able to remember and recite well over 100.000 digits of \(\pi\) \cite{pi-world-ranking-list}, so what are a few 16 byte keys anyway?}.
They need to be stored, and you get a key from every website you login to that supports this protocol.
Before you know it, you have over 50 keys.
\par
All keys will be stored on a digital key ring, which is just a set of 128-bit random numbers.
Each key is associated with a website and automatically selected.
%%%
\Subsubsection{The user identifier}
\begin{dialogue}
\speak{Centurion}	Where is Brian of Nazareth? I have an order for his release.
\speak{Mr. Cheeky}	Uh, I am Brian of Nasareth!
\speak{Brian}		What?
\speak{Mr. Cheeky}	Yeah, I - I - I'm Brian of Nazareth.
\speak{Centurion}	Take him down!
\speak{Brian}	I'm Brian of Nazareth!
\speak{Victim \#1}	Eh, I'm Brian!
\speak{Mr. Big Nose}	I'm Brian!
\speak{Victim \#2}	Look, I'm Brian!
\speak{Brian}		I'm Brian!
\speak{Victims}		I'm Brian!
\speak{Gregory}		I'm Brian, and so's my wife!
\attrib{Monty Python's ``Life of Brian''}
\end{dialogue}
There are several things that make traditional userid's cumbersome to use.
\begin{itemize}
\item Anyone can claim to be `Jessica' and try to login on her behalf.
If her password has been compromised or guessed,
the real Jessica may either have her privacy violated or her reputation destroyed.
If `she' has committed a crime, Jessica has a hard time proving it was not her.
\item They are in short supply (there are many Jessicas in the world, all (perhaps narcistically) preferring to use just 'jessica').
\item Each website has its own set of userids, which may make `jessica77' available on one site, but not on the other.
This leaves Jessica with many different userids, each of which she must remember, or write down.
\item Especially email userids can be used to link accounts between different websites.
Since email addresses are fairly unique, your browsing behaviour may be tracked through logins.
\end{itemize}
Since we are letting the computer join and aid the login process, why not compute some more and come up with a large, fairly random, value that is used to identify a person.
One that is unique; not per person (like mail accounts), but to an account per user per webserver.
\par
Instead of sending a legible string, why not send something that is not relatable to a human: a 256 bit hash, for example.
This hash could be made up of many things, and it has always been the task of a computer to match userids to accounts, so a hash value would do the trick just as well.
\par
The idea is that you cannot guess this value
(only Jessica knows how to generate one)
and that it may be compiled by using several sources,
so this authentication protocol becomes genuinally multi-factor.
%%%
\Subsubsection{Login server}
Designing and maintainig a website, especially large ones, is a specialist's job.
Interoperability with payment systems requires specialists too.
The same goes for the databases running the show.
\par
But securing keys and operating them is as much a specialist's job as any mentioned before, maybe more so.
This is not something you do on the side and hope everything will be all right.
Typically, none of all the specialists you have working on your website will have indepth knowledge of this field of expertise.
\par
As it turns out, all cryptographical operations can be singled out and combined into what is called a Login server.
This also means that operating the Login server could be done by people other than those running the website traditionally
(which can keep their job and do their thing without change).
In fact, the whole business running Login servers can be put on the Internet as LaaS: Login as a Service.
\par
This service can be protected and secured by professionals that know nothing about websites,
but all the more from securing Internet traffic,
protecting keys in HSMs,
do key rollovers,
organize key ceremonies,
generate certificates,
and everything else you need to make the Login servers little bastions of security.

\input{tex/user_hash.sub.tex}

%%%%%%%%%%%%%%%%%%%%%%
\Subsection{A rudimentary login algorithm}
\begin{dialogue}
\speak{Dr. Strange}	I went forward in time, to view alternate futures. To see all the possible outcomes of the coming conflict.
\speak{Star-Lord}   How many did you see?
\speak{Dr. Strange}	14,000,605.
\speak{Iron Man}	How many did we win?
\speak{Dr. Strange}	One.
\attrib{Avengers: Infinity War}
\end{dialogue}
The core idea of the algorithm is that the user must prove it is in the posession of a key that the server has issued.
To do so, there is an almost limitless number of approaches to pursue, but not all will be fruitful.
\par
In the solution we are disussing now, the server encrypts a random number (a nonce) with the key the user has.
The user can prove it has that key by decrypting the nonce and returning a value derived from it, like a hash for instance.
\par
But, we do not want the web server to have all the user keys, or any other directly usable value, for that matter.
This is the main inherent flaw of authentication using passwords.
Instead, we make the user key a direct derivative of the site key.
\par
There are several ways to achieve this; we do it with a calculation similar to
\[K_u = \EAES{K_s}{S}\]
The key the user has (\(K_u\)) is the \AES\ encrypted value (with some key \(S\)) of the key the website has (\(K_s\)).
More of this later.
\par
For symmetry, we ask the webserver to authenticate itself also, by similar means: decrypt the client's random value and send proof you did that with the right key.
\par
What follows is a step by step unrolling of the protocol, where each calculation is described.
%%%
\Subsubsection{Key length \(n\) indicated by the server}
At the start of the login protocol the webserver indicates the key lenght \(n\) to the user.
\[n\quad\Bigl\lvert\thickspace\text{with}\thickspace 1\le{n}\le128\]
This is the length of the keys we are going to use.
Not always the maximum is chosen; a value of 32 bit is a valid and convenient choice also.
%%%
\Subsubsection{Computing challenge \(A_u\) by the user}
We need one of our keys, stored on the key ring, to help us with the calculations.
A lookup is done to get the key for the website.
Suppose we find it at index 47 and compute:
\begin{equation}\label{rud:uku}
K_u = \mtpn{Z[47]}
\end{equation}
where the leftmost (most significant) $128 - n$ bits are stripped with the modulo operation.
This yields a key of precisely $n$ bits, as instigated by the website.
\par
Then, the user generates a random number, also of \(n\) bits, and encrypts it with the key it has for that website:
\begin{equation}\label{rud:uru}
R_u = random(n)
\end{equation}
\begin{equation}\label{rud:au}
A_u = \xor{R_u}{K_u}
\end{equation}
\par
Which makes this an \(n\)-bit value.
We use \(A_u\) as a challenge for the website: please decrypt it and send proof you have found our \(R_u\).
%%%
\Subsubsection{Computing response \(B_u\) by the user}
As proof of having found \(R_u\), we expect a value in return of our \(A_u\), which proves that we are dealing with the right website.
Just sending a 256-bit hash value
\(B_u = \HSHA{R_u}\)
would expose \(K_u\).
With some effort, all n-bit random numbers could be checked and \(K_u\) computed.
\par
By taking just the least \(n\) significant bits of the hash:
\(B_u = \mtpn{\HSHA{R_u}}\)
our random number is now hidden in a large set of \(2^{256-n}\) possible values.
That, in itself, does not hide it completely.
Especially with a large \(n\), there may be few (if any) collisions, so a brute force attack may still be fruitful.
Taking it \(\bmod\ 2^n\) does give us the opportunity to encrypt it with \(K_u\):
\begin{equation}\label{rud:bu}
B_u = \xor{\mtpn{\HSHA{R_u}}}{K_u}
\end{equation}
\par
Possibly many random numbers yield the same hash (taken \(\bmod\ 2^n\)).
This is no problem, because we use this random value only once, ever.
Producing this hash value is ample proof that the website has the right key.
\par
The user keeps the \(B_u\) value in memory, until the server returns its computations.
%%%%
\Subsubsection{Computing \(B_s\), \(P_s\), and \(Q_s\) on the server}
The \(A_u\) value is sent to the server, along with the generated \(H_u\) value that was built in an earlier step.
Upon receiving \(A_u\) the server will try to decrypt it.
For that, it needs the user key \(K_u\), which it can obtain using the site key \(K_s\), which is stored in the accounts database under index \(H_u\):
\begin{equation}
K_s = account\_lookup(H_u)
\end{equation}
\par
So, the next cacluations are needed: first get hold of the user key \(K_u\) and then extract the user random \(R_u\).
\begin{equation}\label{rud:sku}
K_u = \EAES{K_s}{S}
\end{equation}
The key \(S\) is the secret key.
We will be discussing this at length later.
\begin{equation}\label{rud:sru}
R_u = \xor{A_u}{K_u}
\end{equation}
Having obtained the \(R_u\) value, we can now compute the value that the user expect us to return.
Note that the \(B_s\) value is computed exactly like \eqref{rud:bu}.
\begin{equation}\label{rud:bs}
B_s = \xor{\mtpn{\HSHA{R_u}}}{K_u}
\end{equation}
The value \(B_s\) is stored in memory for a short while.
\par
Next, we generate challenge \(P_s\) for the user:
\begin{equation}\label{rud:rs}
R_s = random(n)
\end{equation}
\begin{equation}\label{rud:ps}
P_s = \xor{R_s}{K_u}
\end{equation}
The \(Q_s\) value will be the expected return value, which is computed slightly different than \eqref{rud:bs}, now using \(R_u\).
This makes the protocol asymmetrical by design.
\begin{equation}\label{rud:qs}
Q_s = \xor{\mtpn{\HSHA{R_s}}}{R_u}
\end{equation}
\par
We keep \(Q_s\) for later use; the values \(B_s\) and \(P_s\) are sent to the user and can be erased from memory right thereafter.
%%%
\Subsubsection{Validating \(B_u\) and computing \(Q_u\) by the user}
The user should assert that value \(B_s\) is  matching \(B_u\);
if not, the login has failed.
\par
If they match, we then can compute \(Q_u\) as a response to \(P_s\):
\begin{equation}
R_s = \xor{P_s}{K_u}
\end{equation}
Having extracted \(R_s\) we compute \(Q_u\) just like \eqref{rud:qs}:
\begin{equation}
Q_u = \xor{\mtpn{\HSHA{R_s}}}{R_u}
\end{equation}
This value is returned to the website to complete the authentication process.
%%%
\Subsubsection{Finalizing authentication on the server}
When the user sends a \(Q_u\) that matches \(Q_s\), we have sucessfully completed the authentication process.
The user is now logged in.
