%
% $Header: /usr/local/src/slwp/article/tex/keys.tex,v 1.8 2014/02/01 13:15:42 timo Exp $
%
% AUTHOR:	ing. T.M.C Ruiter
%
\Section{Keys}
The keys used in this login scheme do not all have the same length.
The Secret keys have 256 bits, for instance.
The $\mathtt{RSA}$ operations use keys of 2048 bits.
Keys stored in a key ring have 128 bits,
but fewer bits may actually be used.
The number of bits in such a key may differ among websites,
and is denoted in this article as value $n$.
%See \vref{sec:key_length} for a discussion of key lengths.
\par
Three sorts of keys are used:
user keys
(on a key ring),
site keys
(in a database),
and Secret keys
(in an array in an HSM).
A user key is the result of encrypting the site key for that user with a Secret key.

\Subsection{Secret keys}
\label{sec:secret_keys}
The keys used to encrypt other keys%
---called Secret keys in this article---%
are used to perform \AES\ encryptions of site keys to get user keys,
and are 256 bits long.
They reside in the HSM of a login server.
\subsubsection{Array of Secret keys}
For enhanced security,
Secret keys need to be replaced at regular intervals.
This change of keys is initiated by the website,
without the possibility to make individual arrangements with any of its users.
\par
If there were only one Secret key,
changing it would render all user keys permanently useless.
Therefore,
an array of Secret keys needs to be kept
(with room for at least one old Secret key),
allowing users to login using an old(er) key.
Several Secret keys are used this way and
are considered to be stored in array $S$, in this article.
\begin{table}[h]
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
1 & 2 & \ldots & $\mathtt{MAX\_ACTIVE\_KEYS}$ & \ldots & $\mathtt{MAX\_KEYS}$ \\
\hline
\end{tabular}
\end{table}
\par
With this array, three values are defined.
\begin{enumerate}
\item	The number of stored Secret keys is represented by $m$.
\item	The constant%
\footnote{Although declared a constant, the value of \texttt{MAX\_KEYS} may change over time.
When the login policy regarding the use of old keys is changed, more or fewer old keys will be stored.
For this, \texttt{MAX\_KEYS} needs to be changed.}
\texttt{MAX\_KEYS}, which is the maximum number of keys that will be kept.
The value of $m$ starts at 0 and will reach \texttt{MAX\_KEYS} eventually and grow no more.
The value \texttt{MAX\_KEYS} has a minimum of 2 (a new key and at least 1 old key).
All values in the array $S$ are shifted up one position when a new key is stored, which will always be inserted at $S[0]$.
This way, with at least one key loaded ($m>0$), it always holds that $S[0]$ is the newest key,
and $S[m-1]$ is the oldest key.
With $m>1$, $S[1]$ is the youngest old key.
\item	The constant \texttt{MAX\_ACTIVE\_KEYS},
which is the maximum number of keys that will be used for logging users in.
This value lies in the range $[1, \mathtt{MAX\_KEYS}]$.
\par
Subtracting \texttt{MAX\_ACTIVE\_KEYS} from \texttt{MAX\_KEYS}
gives the number of inactive keys which can be used to reactivate expired user keys.
Users using a user key that is older than the oldest active Secret key cannot login,
but their key can be restored with an inactive Secret key.
If a key is used that is older than the oldest inactive key,
the key cannot be restored and is lost forever.
\end{enumerate}
\par
Upon successful login with an old key,
the user will be provided with a new user key automatically.
%with which it can login using the newest Secret key,
%from that moment on.
The new user key is based on the newest Secret key and can be used for logins as long as that Secret key in use.
Changing of user keys is seamless
(see section~\vref{sec:userkeys})
and invisible to the user.
%and the user might,
%just as well,
%be kept unaware of this.
\par
The website's security policy should determine with what frequency Secret keys are to be changed,
and how many old keys should be kept.
If,
for instance,
the Secret key is changed every month,
and 11 old keys are kept,
users can still login using a key that is up to a year old.
If the user key is older than that,
the site is not able to verify the user's key any longer.

%------------------------------------------------------------------
\subsubsection{Terminated accounts}
When deleting the oldest Secret key from memory,
all accounts with a `latest login' date older than the installation
date of the second oldest Secret key should be marked `terminated'.
Removal of the oldest Secret key renders unusable all user keys associated with it.
Those accounts can be purged, as they cannot be used again.

%------------------------------------------------------------------
\subsubsection{Expired accounts}
In some situations user access may be barred until some condition is met
(see also \vref{sec:conditional_key_replacement}).
Accounts can be made temporarily inaccessible for this purpose by letting keys expire.
Expiration can happen automatically when users do not login in time to get their keys replaced,
or purposely by denying key updates.
Expired user keys are associated with Secret keys with an index in the range $[\texttt{MAX\_ACTIVE\_KEYS},\texttt{MAX\_KEYS})$.
\par
Expired accounts can easily be made active again by using inactive Secret keys for the login process
(and then renew the user key).
\par
Expired accounts can be reinstated,
but only before the associated Secret key is erased from memory.
After that, there is no way to regenerate the user key for logging in.
The user should apply for a new account if he/she wants to regain access.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{Site keys}
For each user,
a site key is generated once by a pseudo-random generator of the HSM of the login server
(see section~\vref{sec:apply_step3}),
and never needs to be replaced.
\par
The site keys are stored in a table of a database of the accounts server,
where a hash value will act as the primary key for that table.
Each time a user tries to login,
the database is queried with the hash value the user supplies,
and the site key for that user is returned for further processing.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\Subsection{User keys and key rings}
\label{sec:userkeys}
Together with the site key,
a user key is computed by encrypting the site key for the user with the youngest
(or current)
Secret key,
using the \AES\ algorithm
(also see section~\vref{sec:apply_step3}).
\par
The keys users get from the website are stored on a key ring (see section~\vref{sec:key ring}).
The key ring will be considered an array $Z$ in the rest of the article.
A key in a key ring is automatically replaced with a new key by the web server
whenever the Secret key of the login server is changed
(see section~\vref{sec:login_step6}).

\subsubsection{Storing user keys in a key ring}
\label{sec:key ring}
A key ring is created by using a random generator of sufficient quality,
generating a set of random numbers of 128 bits,
stored in a file.
Key number zero is used to identify the key ring and will never change after its creation.
The other keys are dummies
(random bits with no meaning whatsoever),
some of which will be overwritten with real keys over time.
Real keys in the key ring should be put randomly between the dummy keys.
\par
Typically,
100 random numbers are generated this way,
so all keys are conveniently numbered with a one or two digit number
(1 through 99),
which can easily be remembered.
However,
you can put any number of keys in a key ring file.
From zero keys
(which leaves only the key ring identifier $Z[0]$,
and allows you to login to zero websites)
up to any number of keys you bother to carry around with you.
\par
Having more keys than sites you want to login to is just a way to make things
a bit harder for those that do not own the key ring to choose keys.
If you are not happy with this,
you can just add new keys as you acquire them,
just as with real keys.

\subsubsection{Copying key rings}
A key ring can be freely copied to other devices, so all keys are available there as well.
\par
Encrypted key rings can be stored in a public place, for easy access, and act as a backup.
A dedicated web server can be employed for storing encrypted key rings,
which can be used to restore lost key rings.
They can also be used to login when away from home with no access to your own key ring.
These temporary key rings should be discarded when logging out.

\subsubsection{Irretrievably lost keys and key rings}
When a key number is forgotten,
or a key inadvertently overwritten,
and no record or backup of it can be found,
the key must be considered lost.
\par
As the user has supplied websites with personal information,
it may be easy to regain login capabilities by just asking for a new key,
provided a username and maybe some other information can be given.
Selecting a free key number on the key ring and replacing that key with the new one should restore the ability to login.
\par
Key rings can become unusable or lost in two situations:
the device holding it is defective
(like a hard disk crash)
and no backup has been made,
or the encryption cannot be reversed
(PIN forgotten).
In both cases,
the user has to start over and create a new key ring,
containing only dummy keys.

\Subsection{Certificates and key pairs}
As part of the security of the login process,
keys need to be exchanged in encrypted form to and from the login server.
Furthermore,
the website may want to notify its users that it is using a trusted login server.
\par
To accommodate this requirement, the login server needs a certificate that holds a public encryption key.
The private key that belongs to it is stored in the HSM.
With this certificate the users of the login system can validate the trustworthiness of the used login server
by validating the chain of trust.
The certificate is signed by a Certificate Authority (CA).
This CA's certificate is also signed, etcetera, up to a Root CA.
A list of all trusted Root CA's is stored in the browser of the user
and is already used to validate the certificates of websites.
%\subsubsection{Login server keys}
\par
The login server uses two types of \RSA\ keys:
an encryption key and a signing key; both private parts tucked away in its HSM.
The public counterparts of these keys are presented to the user with a certificate,
signed by a Certificate Authority so the user can validate them.
\par
The encryption keys of the login server are called $K_{le}$ for the public key
(`l' for `login'; `e' for `encryption'),
and $K_{LE}$ for the private key.
The signing keys of the login server are called $K_{ls}$ for the public key
(`s' for `signing'),
and $K_{LS}$ for the private key.
%\subsubsection{Accounts server keys}
%Like the login server, the accounts server needs its own set of keys for encryption of other keys.
%No signing is necessary,
%so we have $K_{ae}$ as the public key
%(`a' for `account')
%and $K_{AE}$ as the private key.
