\documentclass[a4paper]{article}
\pagestyle{plain}
\setlength{\textwidth}{15cm}
\setlength{\hoffset}{-1cm}
\setcounter{secnumdepth}{-2}
\newcounter{clausenr}
\newcommand{\clause}[1]{\addtocounter{clausenr}{1}\subsubsection{Clause \Roman{clausenr} --- #1}}
\renewcommand{\theenumi}{\alph{enumi}}
\begin{document}
\section{Assessor Non-disclosure Agreement}
\par
\vspace{1.5cm}
T.M.C. Ruiter\ \  B ICT\hfill
(hereinafter referred to as ``Inventor'')\\[.5cm]
on the one part, and\\[.5cm]
\input{assessor-name.tex}\hfill
(hereinafter referred to as ``Assessor'')\\[.5cm]
on the other part,\\[.5cm]
hereby agree that,
except as may otherwise be agreed in writing,
the following terms and conditions shall be applicable to all meetings and communications between
Inventor and/or representatives of Inventor and Assessor,
in connection with the assessment of market and licensing potential of the invention described in
\begin{enumerate}
\item the article ``Secure Login without Passwords'',
or any other document from the Inventor related to this article; and
\item programming code simulating or truly executing logins using the principle described in said article;
\end{enumerate}
hereinafter called ``The Authorized Purpose''.

\clause{Transfer of information}
Information may be transferred between the Inventor and Assessor to the extent necessary to fulfill The Authorized Purpose.
All information,
orally or otherwise communicated,
is to be regarded as ``Confidential'' and is hereinafter referred to as ``Confidential Information''.

\clause{Property}
All Confidential Information disclosed or transferred by the Inventor to the Assessor shall remain the property of the Inventor.

\clause{Use and protection}
The Assessor agrees that,
unless the Inventor gives his prior written authorization,
the Assessor shall,
during a period of five (5) years from the date of receipt of any Confidential Information hereunder:
\begin{enumerate}
\item Not use the Confidential Information disclosed by the Inventor to the Assessor for any other purpose than for The Authorized Purpose; and
\item Protect the Confidential Information against disclosure to third parties,
in the same manner and with the same degree of care,
but not less than a reasonable degree of care,
with which it protects confidential information of its own.
\item limit circulation of the Confidential Information disclosed by the Inventor to such employees of its own,
of their authorized affiliated companies and of authorized third parties as have a need to know in connection with The Authorized Purpose.
\end{enumerate}
The Assessor hereby represents and warrants towards the Inventor that each of its employees
receiving the Confidential Information shall be under a legal obligation to observe secrecy and not to
disclose the Confidential Information.

\clause{Not confidential information}
The Assessor agrees that information disclosed pursuant to this Agreement which would otherwise be
Confidential Information shall not be deemed Confidential Information to the extent that it:
\begin{enumerate}
\item is or becomes part of the public domain without violation of this Agreement; or
\item is known and on record at the Assessor prior to disclosure by the Inventor; or
\item is lawfully obtained by the Assessor from a third party; or
\item is furnished to others by the Inventor without restrictions similar to those herein contained as to the use or disclosure by the Inventor; or
\item is developed by the Assessor completely independently of any such disclosure by the Inventor; or
\item is ascertainable from a publicly available product; or
\item is required by law or court order to be disclosed.
\end{enumerate}

\clause{Reproduction, return, and destruction}
The Assessor shall reproduce Confidential Information only to the extent necessary for The Authorized Purpose.
On written request from the Inventor,
the Assessor shall return,
or have an officer certify the destruction of,
all originals and copies of Confidential Information,
within fourteen (14) days.
Upon written request of the Inventor at any time during this Agreement,
or at most ninety (90) days after termination of this Agreement,
the Assessor undertakes to fully and irretrievably delete any Confidential Information contained
in electronic data processing equipment or electronic storage media
(such as hard disks, tapes, diskettes, CD-ROMs, etc.)
of the Assessor,
within fourteen (14) days.

\clause{Forced disclosure}
In the event the Assessor becomes legally required to disclose any Confidential Information,
such Assessor shall promptly and in advance of making such disclosure notify the Inventor of such legal requirement,
so that the Inventor may,
if he wishes to do so,
seek an appropriate remedy to prevent the disclosure.
The Assessor will,
if requested by the Inventor,
cooperate with the Inventor in seeking such remedy.

\clause{Breaches}
Should this Agreement be breached,
money damages for breach of contract will be due for payment.
Furthermore,
any court of competent jurisdiction may enjoin the Assessor from disclosing or utilizing the Confidential Information.

\clause{No obligation}
This Agreement creates no obligation to purchase,
sell,
develop,
research or disclose anything.
It creates no agency or partnership.
This Agreement furthermore does not represent any commitment by any of the parties to enter into any further agreement with the other party.

\clause{Written communications}
Written communications requesting or transferring Confidential Information under this Agreement
shall be addressed only to the respective designees as follows
(or to such designees as the parties hereto may from time to time designate in writing):

\input{assessor-address.tex}

\clause{Sole purpose}
This Agreement is solely for the purpose of governing the disclosure(s) as defined herein
from Inventor to Assessor in the context of The Authorized Purpose.
In the event that the parties enter into an agreement concerning a business relationship,
the provisions of such agreement concerning confidentiality of information shall supersede and
prevail over any conflicting provisions of this Agreement.

\clause{Entire understanding}
This Agreement constitutes the entire understanding between the parties and may not be modified except by agreement in writing,
signed by all parties.

\clause{Invalid provisions}
In the event that any of the provisions of this Agreement are,
or become,
invalid because they are inconsistent with the applicable law,
this shall in no manner affect the other provisions of the Agreement as well as the Agreement as a whole.
The parties shall be obliged to replace such invalid provisions by new legally valid ones having the most similar economic and/or legal effect.

\clause{Failures}
Failure of any party at any time to require performance of any provisions of this present Agreement
shall not affect the right to require full performance thereof at any time thereafter.

\clause{Governing law}
This Agreement shall be governed by and construed in accordance with the laws of The Netherlands. 
\par
Should any dispute arise amongst the parties,
they shall make their best effort to resolve such a dispute amicably amongst themselves.
Disputes arising in connection with this Agreement,
if not amicably resolved by both parties,
shall be settled by the competent courts at the domicile of the defendant.

\clause{Effectuation and termination}
This agreement shall be effective with retroactive effect if applicable,
\input{assessor-date.tex}
for a period of 60 months as from that date,
provided it has been signed by all parties.
%The obligations pursuant to Clause III will survive the term of this Agreement.

\vspace{1cm}

\begin{tabbing}
XXXXXXXXXXXXXXXXXXXXXXXXXXXX \= \kill
\bfseries Inventor: \> \bfseries Assessor: \\
\\
T.M.C. Ruiter \> \input{assessor-name.tex}\\
\end{tabbing}
\begin{tabbing}
XXXXXXXXXXXXXXXXXXXXXXXXXXXX \= \kill
\makebox[5cm]{\dotfill} \> \makebox[5cm]{\dotfill} \\
\tiny date \> \tiny date
\end{tabbing}
\vspace{1cm}
\begin{tabbing}
XXXXXXXXXXXXXXXXXXXXXXXXXXXX \= \kill
\makebox[5cm]{\dotfill} \> \makebox[5cm]{\dotfill} \\
\tiny signature \> \tiny signature
\end{tabbing}
\end{document}
