\section{Required hardware}
Apart from a server to host the website itself,
you need a Secure Device Provisioning Server (SDPS)
that contains a Hardware Security Module (HSM),
which stores an array of Secret Keys and performs cryptographic functions.
Yet another server should host the accounts database, linking at least userids with site keys.

\subsection{Firewall}
There should be a firewall between the webserver and the SDPS,
to protect the SDPS from attack should the website be compromised.
The SDPS should be placed in a DMZ.
The other link from the firewall will be to the server that hosts the accounts database.

\subsection{Accounts database}
There should be a server for the database with account information.
This database will probably be used for many other things that the website needs as well.

\subsection{SDPS}
\label{sec:SDPS}
The Secure Device Provisioning Server (SDPS) has a Hardware Security Module (HSM)
for storing cryptographic keys, performing cryptographic functions, and generating random numbers.
It should be separated from the rest of the website's internal network by a firewall,
and placed in a Demilitarized Zone (DMZ) to protect it from misuse or attack.
\par
The data that is sent to the SDPS, and the data that is received back from the SDPS are encrypted by design.
Securing this channel with extra encryption is not really needed, as network sniffing yields only random values.
\par
As such, a single SDPS can be shared among websites or multiple hosts for one large website, and its data sent over the Internet.
\par
For redundancy, multiple SDPS systems should be employed, for if an SDPS fails, no one can login.
When changing Secret keys in this scenario
the keys need to be replaced simultaneously.
When a user got a new key from an SDPS with the new key, and logs in again, an SDPS that misses the newest Secret key cannot verify the user any more.

\subsection{HSM}
The Hardware Security Module is a piece of tamper proof hardware.
It stores cryptographic keys, certificates, and other data in a secure manner.
It also has a whealth of cryptographic functions and can generate good quality random numbers.
The HSM should be programmed in such a manner that even the SDPS, which hosts the HSM, is only able to use a small set of functions.
\par
Ideally, the set of functions the SDPS should provide should be programmed directly into the HSM.
This way, the Secret keys can be kept secret, and the SDPS cannot be misused by a hacker to regenerate login data.
