%
% $Header: /usr/local/src/slwp/article/tex/keyring.tex,v 1.5 2013/03/09 15:27:32 timo Exp $
%
% AUTHOR:	ing. T.M.C. Ruiter
%
\section{Storing user keys in a keyring}
\label{sec:keyring}
Keys for the user are collected and stored in a keyring.
The keyring is a block of at least 100 random numbers with 128 bits, most of them dummy keys.

\subsection{Creating a keyring}
A keyring is created by using a random generator of sufficient quality,
which generates 100 random numbers of 128 bits.
Key number zero is used to identify the keyring and will never change after its creation.
The other 99 keys are dummies
(random bits with no meaning whatsoever).
These random numbers are then encrypted and written to a file on a storage device.
Unencrypted values should never be written to a storage device and only be kept in volatile memory.
It is up to the user to remember how to decrypt its keyring.
An encrypted keyring is of no value to a hacker without knowledge how to decrypt it.
(See also section \ref{sec:encrypting_a_keyring}).

\subsection{Copying keys and keyrings}
A keyring can be freely copied to other devices, so all keys are available there as well.
The keys of one keyring can be copied to other keyrings, in different locations.
It is up to the user to keep a registration of this.
\par
Encrypted keyrings can be stored in a public place, for easy access, and act as a backup.
A dedicated webserver can be employed for storing encrypted keyrings,
which can be used to restore lost keyrings.
They can also be used to login when away from home with no access to your own keyring.
These temporary keyrings should be discarded when logging out.

\subsection{Irretrievably lost keys and keyrings}
When a keynumber is forgotten and no record of it can be found,
the key must be considered lost.
\par
As the user has supplied websites with personal information,
it may be easy to regain login capabilities by just asking for a new key,
provided a username and maybe some other information can be given.
Selecting a free keynumber on the keyring and replacing that key with the new one should restore the ability to login.
\par
Keyrings can become unusable or lost in two situations:
the device holding it is defective (like a harddisk crash) and no backup has been made,
or the encryption cannot be reversed (PIN forgotten).
In both cases, the user has to start over and create a new keyring, containing only dummy keys.
